---
title: 'Group#2 Homework #4'
author: "Group 2"
date: "4/8/2018"
output: html_document
---

#Introduction
The purpose of this homework assignment is to explore, analyze and model a dataset containing 8161 observations and 26 variables. The dataset are records representing a customer at an auto insurance company.

Each record has two response variables. The first response variable, TARGET_FLAG which is binary (0,1). If someone was in car crash the value is 1 and if the person was not in a car cash the value is 0. 

The second response variable is TARGET_ATM. If someone was in a car cash the value is 1 and if they did not crash their car the value is greater than 0.

#Objective and Requirements
The objective is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car. 

Ony the variables that are given or variables derived from the variables provided. 

#Approach
The team met to discuss this assignment and an approach to plan to complete the assignment. Each of the 5 team members was assigned tasks. The following tasks were assigned:

*Data Exploration
*Data Preparation
*Build Models
*Select Models

Github was used to manage the project. Using Github helped with version control and ensured each team member had access to the latest version of the project documentation. 
Slack was used to by the team to communicate during the project and for quick access to code and documentation.
```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(psych)
library(GGally)
library(ggplot2)
library(reshape)
library(VIM)
library(mice)
```
#Dataset
For reproducibility of the results, the data was loaded to and accessed from a Github repository. 
```{r Read data, echo=FALSE, message=FALSE, warning=FALSE}
insurance_train <- read.csv("https://raw.githubusercontent.com/621-Group2/HW4/master/insurance_training_data.csv", header=TRUE, sep=",")

#Remove the index from tehd dataset
insurance_train$INDEX <- NULL

```
#Data Exploration
Several of the predictor variables contain missing values and outliers. Imputation will be used for the missing values.
##Missing Values
The majority of variables do not contain missing values. The preictory CAR_AGE (Vechile Age) contain 510 missing values and YOJ(Years on Job) contain 454 missing values.
```{r miss, echo=FALSE, message=FALSE, warning=FALSE}

sapply(insurance_train, function(x) sum(is.na(x)))

```


##Variables
```{r data exploration, echo=FALSE, message=FALSE, warning=FALSE}
variableNames <- c("TARGET_FLAG", "TARGET_AMT", "AGE", "BLUEBOOK", "CAR_AGE", "CAR_TYPE", "CAR_USE", "CLM_FREQ", "EDUCATION", "HOMEKIDS", "HOME_VAL", "INCOME", 
"JOB", "KIDSDRIV", "MSTATUS", "MVR_PTS", "OLDCLAIM", "PARENT1", "RED_CAR", "REVOKED", "SEX", "TIF", "TRAVTIME", "URBANICITY", "YOJ")

definition <- c("Was Car in a crash? 1=YES 0=NO", "If car was in a crash, what was the cost", "Age of Driver", "Value of Vehicle", "Vehicle Age", "Type of Car", "Vehicle Use", "# Claims (Past 5 Years)", "Max Education Level", "# Children at Home", "Home Value", "Income", "Job Category", "# Driving Children", "Marital Status", "Motor Vehicle Record Points", "Total Claims (Past 5 Years)", "Single Parent", "A Red Car", "License Revoked (Past 7 Years)", "Gender", "Time in Force", "Distance to Work", "Home/Work Area", "Years on Job")

variableType <- c("Response", "Response", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor")

dfInsurance_md <- cbind.data.frame (variableNames, definition, variableType)

colnames(dfInsurance_md) <- c("Variable Name", "Definition", "Variable Type") 
knitr::kable(dfInsurance_md)
```
##Descriptive Statistics
Descriptive statisitics was performed for all predictor and response variables to explore the data. 
```{r descriptive statistics, echo=FALSE, message=FALSE, warning=FALSE}
#Calculate mean missing values per variable
insurance_train %>% summarize_all(funs(sum(is.na(.)) / length(.)))

#Use Describe Package to calculate Descriptive Statistic
(InsuranceTrain_des <- describe(insurance, na.rm=TRUE, interp=FALSE, skew=TRUE, ranges=TRUE, trim=.1, type=3, check=TRUE, fast=FALSE, quant=c(.1,.25,.75,.90), IQR=TRUE))
```

##Correlation Analysis
There is high correlation amoung several predictors CLM_FREQ and MVR_PTS; KIDSDRV and AGE; GET_FLAT and TARGET_AMT; AGE and HOMEKIDS. 
```{r correlation, echo=FALSE, message=FALSE, warning=FALSE}
ggcorr(insurance_train)

```
#Analysis of predictors
Each predictor was examed to determine whether transformation is needed.

##KIDSDRIV
The Driving Children variable is highly skewed to the right. There appear to be outliers
```{r KIDSDRIV, echo=FALSE, message=FALSE, warning=FALSE}
#Density & box plots
par(mfrow=c(1,2))
kidDensity <- density(insurance_train$KIDSDRIV)
plot(kidDensity, main="Driving Kids")
polygon(kidDensity, col="red", border="red")

boxplot(insurance_train$KIDSDRIV, main="Boxplot of Driving Kids", col="blue")

```
##AGE
The AGE predictor is normally distributed with high outliers of ages 72, 73, 76, 80 & 81 ane low 16, 17 and 18.
```{r age, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
ageDensity <- density(na.omit(insurance_train$AGE))
plot(ageDensity, main="Age of Driver")
polygon(ageDensity, col="red", border="red")

boxplot(insurance_train$AGE, main="Boxplot Age of Driver", col="blue")

```
##BLUEBOOK
The predictor of car value BLUEBOOK shape is similar to bimodal. There are some outliers a the higher car value level.
```{r bluebook, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
bBookDensity <- density(as.numeric(insurance_train$BLUEBOOK))
plot(bBookDensity, main="Value of Vehicle")
polygon(bBookDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$BLUEBOOK, main="Boxplot Value of Vehicle"), col="blue")

```
##CAR_AGE 
The distribution of the age of the vechicale is normal. There are several outliers with newer and older cars.
```{r car_age, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
car_ageDensity <- density(na.omit(insurance_train$AGE))
plot(car_ageDensity, main="Vehicle Age")
polygon(car_ageDensity, col="red", border="red")

boxplot(insurance_train$AGE, main="Boxplot Vehicle Age", col="blue")

```
##CAR_TYPE 
z_SUV and Minivan are majority of vehicles insured.
```{r car_type, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(insurance_train, aes(x = CAR_TYPE)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("Type of Car") + ylab("Number of Observations")

table(insurance_train$CAR_TYPE)
```
##CAR_USE
The majority of cars are privately used.
```{r car_use, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(insurance_train, aes(x = CAR_USE)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("Vehicle Use") + ylab("Number of Observations")

table(insurance_train$CAR_USE)
```
##CLM_FREQ 
The distribution of claims is multi modal. With the largest number of claims occuring before year 1.
```{r clm_freq, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
clm_freqDensity <- density(as.numeric(insurance_train$CLM_FREQ))
plot(clm_freqDensity, main="# Claims (Past 5 Years)  ")
polygon(clm_freqDensity, col="red", border="red")

boxplot(insurance_train$CLM_FREQ, main="# Claims (Past 5 Years)", col="blue")
```
##EDUCATION
```{r education, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(insurance_train, aes(x = EDUCATION)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("Type of Car") + ylab("Max Education Level")

table(insurance_train$EDUCATION)
```

##HOMEKIDS 
The distribution of HOMEKIDS is multimodal. The majority of customers do not have any children.
```{r homekids, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
homekidsDensity <- density(as.numeric(insurance_train$HOMEKIDS))
plot(homekidsDensity, main="# Children at Home")
polygon(homekidsDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$HOMEKIDS, main="# Children at Home "), col="blue")
```
##HOME_VAL 
The distribution of HOME_VAL is skewed to the left. There are negative values that will require futher exploration.
```{r home_val, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
home_valDensity <- density(as.numeric(insurance_train$HOME_VAL))
plot(home_valDensity, main="Home Value")
polygon(home_valDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$HOME_VAL, main="Home Value"), col="blue")
```
##INCOME 
The distribution INCOME has uni modal and skewed to the left.
```{r income, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
incomeDensity <- density(as.numeric(insurance_train$INCOME))
plot(incomeDensity, main="Income")
polygon(incomeDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$INCOME, main="Income"), col="blue")
```
##JOB
```{r}
ggplot(insurance_train, aes(x = JOB)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("Job Category") + ylab("V")

table(insurance_train$JOB)
```
##MSTATUS 
```{r mstatus, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(insurance_train, aes(x = MSTATUS)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("Marital Status") + ylab("V")

table(insurance_train$MSTATUS)
```

##MVR_PTS  
The distribution of the MVR_PTS is skewed to the left. 
```{r job, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
mvr_ptsDensity <- density(insurance_train$MVR_PTS)
plot(mvr_ptsDensity, main="Motor Vehicle Record Points")
polygon(mvr_ptsDensity, col="red", border="red")

boxplot(insurance_train$MVR_PTS, main="Motor Vehicle Record Points", col="blue")
```
##OLDCLAIM 
The distribution OLDCLAIM is highly skewed to the left.
```{r oldclaim, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
oldclaimDensity <- density(as.numeric(insurance_train$OLDCLAIM))
plot(oldclaimDensity, main="Total Claims (Past 5 Years) ")
polygon(oldclaimDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$OLDCLAIM, main="Total Claims (Past 5 Years) "), col="blue")
```
##PARENT1
The majority of customers are signle parents.
```{r parent1, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(insurance_train, aes(x = PARENT1)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("Single Parent ") + ylab("V")

table(insurance_train$PARENT1)
```
##RED_CAR 
```{r red_car, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(insurance_train, aes(x = RED_CAR)) + 
  geom_bar(fill = "red", width = 0.7) + 
  xlab("A Red Car ") + ylab("V")

table(insurance_train$RED_CAR)
```

##TIF 
The distribution of TIF is skewed to the left with several outliers.
```{r tif, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
tifDensity <- density(as.numeric(insurance_train$TIF))
plot(tifDensity, main="Time in Force")
polygon(tifDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$TIF, main="Time in Force"), col="blue")
```
##TRAVTIME 
The distribution of TRAVTIME is skewed to the left with several outliers.
```{r travtime, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
travtimeDensity <- density(as.numeric(insurance_train$TRAVTIME))
plot(travtimeDensity, main="Distance to Work")
polygon(travtimeDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$TRAVTIME, main="Distance to Work"), col="blue")
```
##YOJ 
The YOJ distribution is close to normally distributed. There are outliers at both the lower and upper ends.
```{r yoj, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
yojDensity <- density(na.omit(insurance_train$YOJ))
plot(yojDensity, main="Years on Job")
polygon(yojDensity, col="red", border="red")

boxplot(as.numeric(insurance_train$YOJ, main="Years on Job"), col="blue")
```

#Data Preparation
##Missing Values
The majority of cases are complete. Of concern are the 2 predictor variables (CAR_AGE, YOJ) that have more than 5% of missing values. However, the majority of variables have less than 10 missing values.

Predictors without missing values that contain zero values are possible indication zero values are actually missing values. For instance, predictors HOME_VAL and INCOME have zero values which are highly unlikely
```{r missing values, echo=FALSE, message=FALSE, warning=FALSE}
#Number of missing
aggr(insurance_train, prop = T, numbers = T)
```
##Impute Missing Values
The missing data patterns show that 7,213 out of 8,161 are complete observations, 6 observations are missing the AGE predictor, 432 observations are missing YOJ, 488 observations are missing CAR_AGE and 22 observations are missing YOJ and CAR_AGE. 
```{r impute, echo=FALSE, message=FALSE, warning=FALSE}
#Examine missing data pattern
md.pattern(insurance_train)


```


#Transformations
```{r transformations, echo=FALSE, message=FALSE, warning=FALSE}

```

